\textbf{Advection versus diffusion in intraventricular transport}

### 2025-03-26 paper/main.tex
stability discussion

To verify the model implementation with the zebrafish brain ventricles mesh,
we compared velocity norms for varying degrees of refinement of the brain ventricles
mesh~(\Cref{tab:ventricles_norms_BDM}).
We used the original mesh obtained by meshing the surface geometry using
fTetWild~\cite{fTetWild}; the region around the photoconversion region is not
refined on the mesh used for the convergence study.
For the cilia traction vector $\btau = \tau\lambda(\xx) P_{\nn}(\rr)$ introduced in~\Cref{subsec:appendixA1},
we set a constant $\tau = 6.5\times 10^{-4}$ Pa and
$\lambda(\xx)=1$ in the numerical experiments.
We set $\rr = (1, 0, 1)$ in the ventral middle ventricle cilia region
and $\rr = -(1, 0, 1)$ in the dorsal middle ventricle and the anterior ventricle cilia
regions. We used $\gamma=10$ for the BDM penalty parameter.

@@

The $\mathrm{BDM}_1$-$\mathrm{DG}_0$ scheme that we employ for the Stokes equations
is computationally more expensive than a more conventional approach based on
e.g. Taylor-Hood discretization~\cite{Stenberg1990ErrorProblem}
($P_2-P_1$ elements) or lower-order enriched
elements~\cite{Brezzi2011MixedMethods},
such as the mini-element ($P_1-P_1$ with a velocity stabilization).
We opted for the BDM-DG scheme because of its exactly-mass-conserving property,
since numerical experiments proved that a divergence-free velocity field was
fundamental to attain a stable discretization of the advection-diffusion equation
in advection-dominated regions. Preliminarily, we attempted
to discretize the Stokes equations with 
Taylor-Hood elements, but with the resulting numerical approximation
of the velocity $\uu_h$, we were unable to achieve
stability of the advection-diffusion discretization.
Numerical experiments identified non-zero divergence of $\uu_h$
as the root cause of these stability issues.
On the standard computational mesh, the $L^2$ norm of $\nabla\cdot\uu_h$ was
in the order of $10^{-6}$, while the
maximum value of the divergence of the velocity was of order unity.
On a refined mesh with roughly six times as many
computational cells as the standard mesh, the numerical values of
$\normltwo{\nabla\cdot\uu_h}$ and $\normlinf{\nabla\cdot\uu_h}$
remained practically unchanged. This is a result
of the fact that incompressibility is only achieved in a weak sense when
employing Taylor-Hood elements. Since the advective flux
\begin{equation*}
    \nabla\cdot(c\uu) = c(\nabla\cdot\uu) + \uu\cdot\nabla c,
\end{equation*}
where in this case $\nabla\cdot\uu\neq 0$, a source (or sink) term for the concentration
is introduced. This may introduce errors in the numerical approximation of $c$
and lead to instabilities and spurious oscillations, especially in the case of
advection-dominated transport
\cite{gresho1998incompressible, cesmelioglu2022compatible, johnson2009numerical}.

>>>>

To verify the model implementation with the zebrafish brain ventricles mesh,
we compared velocity norms for varying degrees of refinement of the brain ventricles
mesh~(\Cref{tab:ventricles_norms_BDM}).
We used the original mesh obtained by meshing the surface geometry using
fTetWild~\cite{Hu2020FastWild}; the region around the photoconversion region is not
refined on the mesh used for the convergence study.
For the cilia traction vector $\btau = \tau\lambda(\xx) P_{\nn}(\rr)$ introduced in~\Cref{subsec:appendixA1},
we set a constant $\tau = 6.5\times 10^{-4}$ Pa and
$\lambda(\xx)=1$ in the numerical experiments.
We set $\rr = (1, 0, 1)$ in the ventral middle ventricle cilia region
and $\rr = -(1, 0, 1)$ in the dorsal middle ventricle and the anterior ventricle cilia
regions. We used $\gamma=10$ for the BDM penalty parameter.

@@

The $\mathrm{BDM}_1$-$\mathrm{DG}_0$ scheme that we employ for the Stokes equations
is computationally more expensive than a more conventional approach based on
e.g. Taylor-Hood discretization~\cite{Stenberg1990ErrorProblem}
($P_2-P_1$ elements) or lower-order enriched
elements~\cite{Brezzi2011MixedMethods},
such as the mini-element ($P_1-P_1$ with a velocity stabilization).
We opted for the BDM-DG scheme because of its exactly-mass-conserving property,
since numerical experiments proved that a divergence-free velocity field was
fundamental to attain a stable discretization of the advection-diffusion equation
in advection-dominated regions. Preliminarily, we attempted
to discretize the Stokes equations with 
Taylor-Hood elements, but with the resulting numerical approximation
of the velocity $\uu_h$, we were unable to achieve
stability of the advection-diffusion discretization.
Numerical experiments identified non-zero divergence of $\uu_h$
as the root cause of these stability issues.
On the standard computational mesh, the $L^2$ norm of $\nabla\cdot\uu_h$ was
in the order of $10^{-6}$, while the
maximum value of the divergence of the velocity was of order unity.
On a refined mesh with roughly six times as many
computational cells as the standard mesh, the numerical values of
$\normltwo{\nabla\cdot\uu_h}$ and $\normlinf{\nabla\cdot\uu_h}$
remained practically unchanged. This is a result
of the fact that incompressibility is only achieved in a weak sense when
employing Taylor-Hood elements. The advective part of the total flux
is
\begin{equation*}
    \nabla\cdot(c\uu) = c(\nabla\cdot\uu) + \uu\cdot\nabla c.
\end{equation*}
In the case that $\nabla\cdot\uu\neq 0$, a source (or sink) term for the concentration
is introduced. This may introduce errors in the numerical approximation of $c$
and lead to instabilities and spurious oscillations, especially in the case of
advection-dominated transport
\cite{gresho1998incompressible, cesmelioglu2022compatible, johnson2009numerical}.
The $\mathrm{BDM}_1$-$\mathrm{DG}_0$ scheme avoids these issues,
since $\nabla\cdot\uu_h=0$ is satisfied exactly as a result of the element definitions.
